Early Observations of the Type Ia Supernova \abc:\\ \textbf{a Case of Interaction with Nearby, Unbound Material and/or} Strong Ejecta Mixing

### 2017-11-16 doc/ms.tex
include revisions to the intro in the manuscript

%% Keywords should appear after the \end{abstract} command. 
%% See the online documentation for the full list of available subject
%% keywords and the rules for their use.
\keywords{methods: observational --- surveys --- supernovae: general --- supernovae: individual (\abc)}

@@

The vast majority of SNe Ia are observed to be powered purely by the
radioactive decay of $^{56}$Ni. While the detection of SN shock cooling or
ejecta interaction is rare, the level of $^{56}$Ni mixing in the SN ejecta can
fundamentally alter the appearance of the SN shortly after explosion
\textbf{(e.g.,
\citealt{2014MNRAS.441..532D,2016ApJ...826...96P,2017MNRAS.466.3784S,2017MNRAS.472.2787N})}.

SNe Ia experience a dark phase after the SN shock breakout but before
radioactive energy diffuses into the photosphere \citep{2014ApJ...784...85P}.
The duration of this dark phase is set by how the newly synthesized $^{56}$Ni
is mixed and deposited into different layers of the ejecta. In the case of
strong mixing, the dark phase is short, or non-existent, as $\gamma$-rays from
the radioactive $^{56}$Ni diffuse to the photosphere rapidly. This also leads
to blue optical colors and a rapid initial rise in the observed light curve.
\textbf{Even with vigorous mixing it is difficult at very early times, $<
1$\,d after explosion, to explain very large luminosities or blue colors,
because the $^{56}$Ni has not had sufficient time to radioactively decay to
$^{56}$Co.} If the mixing is weak and the $^{56}$Ni is confined to the
innermost layers of the ejecta, the dark phase can last for several days. The
early evolution of such SNe results in redder colors and a more moderate rise
in luminosity \citep{2014MNRAS.441..532D,2016ApJ...826...96P}. Thus, the early
light curves of even non-exotic Type Ia SNe convey information about their
progenitor systems by constraining the distribution of synthesized $^{56}$Ni,
which in turn constrains the explosion mechanism.

% \footnote{\citet{2015A&A...574A..60T} found that a significant dark
% phase is not prevalent for a sample of Type Ib/c SNe.}

@@

In this paper, we report observations of an extraordinarily young SN Ia, \abc, which was discovered by the intermediate Palomar Transient Factory (iPTF) on 2016 April $3.36$ UTC at $\textrm{R.A.}=13^h34^m45.49^s$, $\textrm{Dec.}=+13\degr51\arcmin14\farcs3$ (J2000) with a $g_\mathrm{PTF}$-band magnitude of $21.31\pm0.27$ \citep{2016ATel.8907....1M}. The transient is spatially coincident with a tidal tail of the galaxy NGC\,5221, which lies at a distance of $\sim$100\,Mpc. \abc\ is not detected to a limit of $g=22.1$\,mag on April $2.42$, less than 1\,d prior to discovery, and rose by $\sim$2 mag in the 24 hr following its initial detection. Our spectroscopic follow-up
campaign classified \abc\ as a normal SN Ia
\citep{2016ATel.8909....1C}. Our observations and analysis show that the early evolution of \abc\ exhibited several distinct properties relative to SN\,2011fe. We interpret those differences as arising from either strong $^{56}$Ni mixing or ejecta interaction with diffuse material, or a combination of the two.
Alongside this paper, we have released our open source analysis and all of the data utilized in this study. These are available online at \texttt{GitHub} \url{https://github.com/adamamiller/iPTF16abc}.

@@

This paper utilizes the LCO infrastructure for rapid and regular monitoring of SNe, and we thank S.~Valenti and I.~Arcavi for their development efforts.

>>>>

%% Keywords should appear after the \end{abstract} command. 
%% See the online documentation for the full list of available subject
%% keywords and the rules for their use.
\keywords{methods: observational --- surveys --- supernovae: general --- supernovae: individual (\abc, SN\,2011fe)}

@@

The vast majority of SNe Ia are observed to be powered purely by the
radioactive decay of $^{56}$Ni. While the detection of SN shock cooling or
ejecta interaction is rare, the level of $^{56}$Ni mixing in the SN ejecta can
fundamentally alter the appearance of the SN shortly after explosion
\textbf{(e.g.,
\citealt{2014MNRAS.441..532D,2016ApJ...826...96P,2017MNRAS.472.2787N})}.

SNe Ia experience a dark phase after the SN shock breakout but before
radioactive energy diffuses into the photosphere \citep{2014ApJ...784...85P}.
The duration of this dark phase is set by how the newly synthesized $^{56}$Ni
is mixed and deposited into different layers of the ejecta. \textbf{Strong
mixing leads to a short, or non-existent, dark phase, because the radioactive
$\gamma$-rays rapidly diffuse to the photosphere.} This also leads to
\textbf{larger luminosities and bluer optical colors at early times}.
\textbf{Even with vigorous mixing it is difficult at very early times, $<
1$\,d after explosion, to explain very large luminosities or blue colors,
because the $^{56}$Ni has not had sufficient time to radioactively decay to
$^{56}$Co.} If the mixing is weak and the $^{56}$Ni is confined to the
innermost layers of the ejecta, the dark phase can last for several days.
\textbf{Weak mixing} results in redder colors and a more moderate rise in
luminosity \citep{2014MNRAS.441..532D,2016ApJ...826...96P}. Thus, the early
light curves of even non-exotic Type Ia SNe convey information about their
progenitor systems and explosion mechanisms by constraining the distribution
of $^{56}$Ni.
% \footnote{\citet{2015A&A...574A..60T} found that a significant dark
% phase is not prevalent for a sample of Type Ib/c SNe.}

@@

In this paper, we report observations of an extraordinarily young SN Ia, \abc,
which was discovered by the intermediate Palomar Transient Factory (iPTF) on
2016 April $3.36$ UTC at $\textrm{R.A.}=13^h34^m45.49^s$,
$\textrm{Dec.}=+13\degr51\arcmin14\farcs3$ (J2000) with a
$g_\mathrm{PTF}$-band magnitude of $21.31\pm0.27$ \citep{2016ATel.8907....1M}.
The transient is spatially coincident with a tidal tail of the galaxy
NGC\,5221, which lies at a distance of $\sim$100\,Mpc. \abc\ is not detected
to a limit of $g_\mathrm{PTF}=22.1$\,mag on April $2.42$, less than 1\,d prior
to discovery, and rose by $\sim$2\,mag in the 24\,hr following its initial
detection. Our spectroscopic follow-up campaign classified \abc\ as a normal
SN Ia \citep{2016ATel.8909....1C}. Our observations and analysis show that the
early evolution of \abc\ exhibited several distinct properties relative to
SN\,2011fe. We interpret those differences as arising from either strong
$^{56}$Ni mixing or ejecta interaction with \textbf{nearby, unbound} material,
or a combination of the two. Alongside this paper, we have released our open
source analysis and all of the data utilized in this study. These are
available online at \texttt{GitHub}
\url{https://github.com/adamamiller/iPTF16abc}.

@@

This paper utilizes the LCO infrastructure for rapid and regular monitoring of SNe, and we thank S.~Valenti and I.~Arcavi for their development efforts. \textbf{Figure~\ref{fig:branch_vel} would not have been possible without S.~Blondin generously sharing the data from \citet{2012AJ....143..126B}.}
